---
title: "Data Visualization"
author: "Your name"
date: "`r Sys.Date()`"
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

## Load tidyverse

```{r cars}
library(tidyverse)
```

## Iris data

```{r}
# load data
data(iris) 

# Transform data.frame to tibble
iris <- iris %>% as_tibble() 

# View data
glimpse(iris)
```

## Scatter plot

### data + mapping + geom

Use `+` to combine elements.

```{r}
iris %>% 
  ggplot(aes(x=Sepal.Length, y=Sepal.Width,
             colour = Species))+
  geom_point()
```

```{r}
iris %>% 
  ggplot(aes(x=Sepal.Length, y=Sepal.Width,
             colour = Species, 
             size=Petal.Length))+
  geom_point()
```

### data + mapping + geom + scale

Use `+` to combine elements.

```{r}
# scale_color_manual(values =c('blue','red','yellow'))

```


### data + mapping + geom + stats

```{r}
# stat_smooth()

# An equivalent way to produce the same figure:
# every component has its own mapping relationship
```


### data + mapping + geom + stats + coordinate

```{r}
# Sub region with coord_cartesian(xlim = c(5, 7),ylim=c(2,3))

```

### Add description

```{r}
# ggtitle
# labs
# theme(text = element_text(size=14),
#       legend.position="bottom",
#       axis.text.x = element_text(angle = -10) )



```

### data + mapping + geom + stats + facet

```{r}
# facet_grid(~Species)
# multi-panel seperated by Species

```

### Restore and save figures

```{r}
#ggsave(p1,filename = "p1.png",width = 7,height=5)


```

## Box plot

```{r}
iris %>% 
  ggplot(aes(x=Species, y=Sepal.Width))+
  geom_boxplot() +
  coord_flip()
```


## Histogram

```{r}
iris %>% 
  ggplot(aes(x=Sepal.Width))+
  geom_histogram()

# Histogram + density stats
iris %>% 
  ggplot(aes(x=Sepal.Width,y=stat(density)))+
  geom_histogram()+
  stat_density(geom='line',position='identity')
```


## Density plot

```{r}
iris %>% 
  ggplot(aes(x=Sepal.Width,
             fill=Species,color=Species))+
  geom_density(alpha=0.1)
```


## Bar chart

```{r}
iris %>% 
  ggplot(aes(x=Species))+
  geom_bar()

# Compute the percentage of samples with Sepal.Length > 5 for each species
# List of colors: http://sape.inf.usi.ch/quick-reference/ggplot2/colour
# Use the scales package for labeling with percentage
library(scales)

iris %>% 
  mutate(LargeSL=Sepal.Length>5) %>%
  group_by(Species) %>%
  summarize(prop=mean(LargeSL)) %>%
  ggplot(aes(x=Species,y=prop))+
  geom_bar(stat="identity",fill="coral4",alpha=0.8)+
  geom_text(aes(label=scales::percent(prop,
                                      accuracy=0.1,
                                      suffix="%") ),  
            size=3.5,vjust=-0.25,
            position = position_dodge(width = 0.9))+
  ggtitle("Percentage of samples with Sepal.Length > 5 for each species")+
  scale_y_continuous(labels = scales::percent, limits = c(0,1))+
  labs(x="Species",y="Proportion")+
  #coord_flip()+
  theme_bw()

```


## Arranging graphs into a grid with the R package `cowplot`

```{r}
library(cowplot)
p1 <- iris %>% 
  ggplot(aes(x=Sepal.Length, y=Sepal.Width,
             colour = Species))+
  geom_point()+
  stat_smooth()+
  ggtitle("Width vs Length (Sepal)")+ 
  labs(x="sepal length",y="sepal width")+
  theme(legend.position="bottom",
        axis.text.x = element_text(angle = -10) )

p2 <- iris %>% 
  ggplot(aes(x=Petal.Length, y=Petal.Width,
             colour = Species))+
  geom_point()+
  stat_smooth()+
  ggtitle("Width vs Length (Petal)")+ 
  labs(x="petal length",y="petal width")+
  theme(legend.position="bottom",
        axis.text.x = element_text(angle = -10) )

plot_grid(p1, p2)

```

## Pairwise plot with `GGally`

```{r}
library(GGally)
iris %>% ggpairs(aes(colour = Species, alpha = 0.4))

iris %>% ggpairs(aes(colour = Species,alpha = 0.4), columns=1:4)
```

## Interactive plots with `plotly`

```{r}
library(plotly)
p1 <- iris %>% 
  ggplot(aes(x=Sepal.Length, y=Sepal.Width,
             colour = Species))+
  geom_point()+
  stat_smooth()+
  ggtitle("Width vs Length (Sepal)")+ 
  labs(x="sepal length",y="sepal width")+
  theme(text = element_text(size=14),
        legend.position="bottom",
        axis.text.x = element_text(angle = -10) )

ggplotly(p1)
```

## Correlation plot with `corrplot`

```{r}
library(corrplot)

iris %>% 
  filter(Species=="versicolor") %>%
  select(-Species) %>%
  cor(use="complete.obs") %>%
  corrplot.mixed(main="Correlations among variables",
                 mar = c(0, 0, 2, 0))
```

# Others

The package `ggmap` provides a spatial visualization with ggplot2.
The package `gganimate` enable animation generation with ggplot2.

```{r}
library(ggmap)
# Read the base map of London
map_london <- readRDS("map_london_full_July26.rds")
london_lat_lon <- read_csv("Neighbourhood_lat_lon.csv")
london_lat_lon

# simulate some counts
london_lat_lon <- london_lat_lon %>% 
  mutate(fakepop=10+rpois(41,lambda = 1))

london_lat_lon

# Plot the map
ggmap(map_london)

ggmap(map_london) +
  coord_fixed(xlim= -81.245727+c(-0.15,0.15), 
              ylim = 42.986932+c(-0.13,0.13))

# Add labels from london_lat_lon
ggmap(map_london) +
  coord_fixed(xlim= -81.245727+c(-0.15,0.15), 
              ylim = 42.986932+c(-0.13,0.13))+
  # geom_text(data = london_lat_lon, 
  #           aes(x = lon, y = lat, label=fakepop, 
  #               color=cut(fakepop,10)), 
  #           size=5,fontface = "bold", 
  #           show.legend = FALSE)+
  geom_text(data = london_lat_lon, 
            aes(x = lon, y = lat+0.005, 
                label=neigh,
                color=cut(fakepop,10)), 
            size=3.5,  fontface = "bold", show.legend = FALSE)+
  theme(text = element_text(color = "#444444"),
        plot.title = element_text(size = 15, face = 'bold'),
        plot.subtitle = element_text(size = 12),
        axis.text = element_blank(),
        axis.title = element_blank(),
        axis.ticks = element_blank()
  )

```



# Errors in package references

If you loaded the package `MASS` after the pacakge `tidyverse`,
there would be a error when you used the function `select`.

The reason is that `MASS` also has a function called `select`.
 
Restart the R session and run the following again:
```{r,eval=FALSE}
library(tidyverse)
library(MASS)

iris %>% 
  filter(Species=="versicolor") %>%
  select(-Species)

# To fix it, we can explicitly tell R where to find the function
iris %>% 
  filter(Species=="versicolor") %>%
  dplyr::select(-Species) 

```

