---
title: "Data Transfromation"
author: "Your name"
date: "`r Sys.Date()`"
output: 
  html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

## Load tidyverse

```{r cars}
library(tidyverse)
```

## Pipe operator

```{r}
c(1,2,9,4,5,7,8) %>% mean()  
mean(c(1,2,9,4,5,7,8)) #equivalent
```

## Iris data

```{r}
# load data
data(iris) 
iris %>% View()

iris # lengthy output
iris %>% as_tibble() #tibble, simple Data Frames
iris %>%
  as_tibble() %>%
  print(n=15)

# Object class
iris %>% class()
iris %>%
  as_tibble() %>%
  class()

# Update data
iris <- iris %>% as_tibble() 
iris

# View data
iris %>% glimpse()

str(iris)
```


## Subset by columns with `select`

```{r}
?select
iris %>% names()

# Select the first two variables
iris %>% select(Sepal.Length,Sepal.Width) #by name
iris %>% select(1,2) #by index

# Select all but one variable 
iris %>% select(-Sepal.Length)

# Select columns of which names contain "Length"
iris %>% select(contains('Len'))

# Only want numerical variables
iris %>% select_if(is.numeric)

# A problem occur if one directly apply a function to tibble
mean(as.data.frame(iris)[,1])

iris %>%
  select(Sepal.Length) %>% 
  mean()

iris$Sepal.Length %>% mean()

# Extract a column as a vector with pull
iris %>% pull(Sepal.Length) %>% mean() 

# Or use summarize
iris %>%
  select(Sepal.Length) %>% 
  summarise_all(mean)

# To save the result, you need to assign it to a (new) object
irisSub1 <- iris %>%
  select(Sepal.Length,Sepal.Width)

```


## Subset by rows with `filter`

```{r}
?filter

# Keep rows by one condition
iris %>%
  filter(Species=="virginica") 

# Keep rows by two conditions
iris %>%
  filter(Species=="virginica", 
         Sepal.Length>7) #by two variables

# Remove rows with missing Sepal.Length though there is no missing value
iris %>% 
  filter(!is.na(Sepal.Length))

# Remove rows with missing values thought there is no
iris %>% drop_na()

# Find the average of Petal.Length for the Species of virginica
iris %>%
  filter(Species=="virginica") %>%
  pull(Sepal.Length) %>%
  mean()

```


## Sort data with `arrange`

```{r}
?arrange

iris %>% arrange(Sepal.Length)

# descending order
iris %>% arrange(desc(Sepal.Length)) 

# sort by the orders of multiple variables
iris %>% arrange(Sepal.Length, Sepal.Width)
```


## Create new variables with `mutate`

```{r}
?mutate

iris$Sepal.diff <- iris$Sepal.Length-iris$Sepal.Width

iris <- iris %>% 
  mutate(Sepal.LW=Sepal.Length/Sepal.Width,
         Sepal.Area=Sepal.Length*Sepal.Width)

iris %>% print(width=100)

# Overwrite the old variable
iris <- iris %>% 
  mutate(Sepal.Width=Sepal.Width+1)

# The order of new variables matters!
iris %>%
  mutate(SepalLW=Sepal.Length/Sepal.Width,
         PetalLW=Petal.Length/Petal.Width,
         DiffLW=SepalLW/PetalLW)
```

## Compute summary statistics for each column with `summarise_all`

```{r}
?summarise_all

# Different ways of computing the column means
iris %>%
  select(-Species) %>%
  colMeans()

iris %>%
  select(-Species) %>%
  apply(2,mean)

iris %>%
  select(-Species) %>%
  summarise_all(mean)

# Count the missing values
iris %>%
  summarize_all( ~list(sum(is.na(.))/n()))

# ~list() is called Lambda/Anonymous function
# ~list(sum(is.na(.))/n()) is equivalent to
# function(x){sum(is.na(x)/length(x))}
iris %>%
  summarize_all( function(x){sum(is.na(x)/length(x))} )

# Gather columns into key-value pairs
iris %>%
  summarize_all( ~list(sum(is.na(.))/n())) %>%
  gather(key = "variable",value="missingRate")

# Count the unique values for each column
iris %>%
  summarize_all(n_distinct)

```


## Compute summary statistics with `summarize`

```{r}
?summarise

iris %>%
  summarise(m1=mean(Sepal.Length),
            v1=var(Sepal.Length))

iris %>%
  mutate(m1=mean(Sepal.Length),
            v1=var(Sepal.Length)) %>%
  print(width=100)
```

## Compute summary statistics for each group with `summarize` and `group_by`

Notice that the group attribute will be resolved after a summary function is applied.

```{r}
# The base R way of computing summation of each variable within each group
iris %>%
  select(-Species) %>%
  aggregate(by=list(Species=iris$Species),FUN=sum)

# Add the group attribute so that 
# tidyverse knows a groupwise summary statistics should be computed
iris %>%
  group_by(Species)

iris %>%
  group_by(Species) %>%
  summarise_all(mean)

iris %>%
  group_by(Species) %>%
  summarize(m1=mean(Sepal.Length),
            m2=mean(Sepal.Width),
            count=n()) %>%
  print.data.frame()
```

## Rowise group with 'rowwise'

Sometimes, we may need to perform complex operation that mutate is behaving strangely with.
For example, if one wants to find 
the maximum between sepal (petal) length and sepal (petal) width for each sample, 
then we can obtain the result by

```{r}
iris %>%
  select(-Species) %>%
  rowwise()

# Notice the group attribute still exists
iris %>%
  select(-Species) %>%
  rowwise() %>%
  mutate(Sepal.Max=max(Sepal.Length,Sepal.Width),
         Petal.Max=max(Petal.Length,Petal.Width)) %>%
  select(Sepal.Length,Sepal.Width,Sepal.Max)


# Use ungroup() to resolve it
iris %>%
  select(-Species) %>%
  rowwise() %>%
  mutate(Sepal.Max=max(Sepal.Length,Sepal.Width),
         Petal.Max=max(Petal.Length,Petal.Width)) %>%
  ungroup()


```


## Group by multiple variables

```{r}
iris2 <- iris %>%   
  mutate(longSepal=ifelse(Sepal.Length>5.8,"Y","N"))

# group attribute has two layers
iris2 %>% 
  group_by(Species,longSepal)

# One summarize unfold one group layer
iris2 %>% 
  group_by(Species,longSepal) %>%
  summarize_all(mean) 

iris2 %>% 
  group_by(Species,longSepal) %>%
  summarize_all(mean) %>%
  ungroup()

iris2 %>% 
  group_by(Species,longSepal) %>%
  summarize_all(mean) %>%
  select(-longSepal) %>%
  summarize_all(mean)
```


## Join the summary results back to the data with `left_join`

```{r}
irisSummary <- iris %>%
  group_by(Species) %>%
  summarise_all(mean) %>%
  rename(meanSL=Sepal.Length,
         meanSW=Sepal.Width,
         meanPL=Petal.Length,
         meanPW=Petal.Width)

# Traditional way of renaming columns
#names(irisSummary) <- c("Species", "meanSL", "meanSW", "meanPL", "meanPW")

iris %>% 
  left_join(irisSummary, by="Species") %>%
  print(n=10, width=100)
```


