---
title: "Exploratory Data Analysis on Titanic Surivival"
author: "Your name"
date: "`r Sys.Date()`"
output:
  html_document:
    code_folding: hide
    fig_caption: yes
    number_sections: yes
    theme: cosmo
    toc: yes
    toc_depth: 5
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

# Preparations {.tabset .tabset-fade}

## Load libraries

```{r, message=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(scales)
setwd("C:/Users/Yi/Dropbox/newtermphd/Term12/Rworkshop_data_manipulation/exercise/Titanic") 
```


## Read data
```{r data}
titanic <- read_csv("titanic.csv")
```

# Glance at data  {.tabset .tabset-fade}

## Data structure
```{r data str,result='asis', echo=FALSE}
titanic
titanic %>% glimpse()
```


## Number of distinct values for each variable

Make use of `n_distinct`

```{r}
# Count by summarize_all

```


## Missing rate for each variable

```{r}
# Count by summarize_all


# Create a bar chart to display it

```

## Create a variable SurvivedLabel as a factor of Survival

```{r}
titanic

# Create a variable SurvivedLabel as a factor of Survival
titanic <- titanic %>%
  mutate(SurvivedLabel=factor(Survived))

```



# Survival for each variable  {.tabset .tabset-fade}

## Overall

Count the number and proportion of survived passengers.

```{r}
# Step 1: group_by SurvivedLabel
# Step 2: summarise count using count=n()
# Step 3: mutate to find the proportion, count/sum(count)


```


## Pclass

Ticket class.

```{r}
# Use summarise and n() to count each Pclass
# as well as mean() to compute the survival rate
# Step 1: group_by
# Step 2: summarise
titanic %>% 
  group_by(Pclass) %>%
  summarise(sur=mean(Survived), 
            count=n())

# Bar chart for the surival rate
titanic %>% 
  group_by(Pclass) %>%
  summarise(sur=mean(Survived)) %>%
  ggplot(aes(Pclass,sur)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=scales::percent(sur,
                                      accuracy=0.1,
                                      suffix="%") ),  
            size=3.5,vjust=-0.25,
            position = position_dodge(width = 0.9))+
  scale_y_continuous(labels = scales::percent, limits = c(0,1))+
  labs(y="Survival Rate")+
  theme_bw()

```


## Sex

```{r}
# Use summarise and n() to count each Sex
# as well as mean() to compute the survival rate
# Step 1: group_by
# Step 2: summarise


# Bar chart for the surival rate


```

## Age

Use `SurvivedLabel` as a group indicator in boxplot.
Notice that there are many missing values.

```{r}
# Create a histogram of Age


# Boxplot between SurvivedLabel and Age

```

## SibSp

Number of siblings / spouses aboard the Titanic.
The survival rate within each group indicate that 
the survival rate drops for people with higher SibSp except for people who had no siblings.

```{r}
# Use summarise and n() to count each SibSp
# as well as mean() to compute the survival rate
# Step 1: group_by
# Step 2: summarise


```


## Parch

Number of parents / children aboard the Titanic.

```{r}
# Use summarise and n() to count each Parch
# as well as mean() to compute the survival rate
# Step 1: group_by
# Step 2: summarise


```


## Fare

Passenger fare.

```{r}
# Boxplot between SurvivedLabel and Fare


# Boxplot between factor(Pclass) and Fare


```

## Embarked

Port of Embarkation.

```{r}
# Use summarise and n() to count each Embarked
# as well as mean() to compute the survival rate
# Step 1: group_by
# Step 2: summarise


# Bar chart for the surival rate


```


## Sex and Pclass

```{r}
# Survival rates and counts for the combinations of Sex and Pclass
# Use mutate(SexPclass=paste0(Sex,"_",Pclass))
# to create an interaction variable


# Bar chart for the surival rate for each combination


```

# Others {.tabset .tabset-fade}

## Correlation

Find Spearman Rank's correlations among
Survived, Pclass, Age, SibSp, Parch, Fare.
Hint: Use `cor(use="complete.obs", method = "spearman")`.

```{r}
library(corrplot)

# Step 1: Use select to keep the columns: Survived, Pclass, Age, SibSp, Parch, Fare
# Correlations on all the passengers


# Step 1: Use filter to keep observations with Survivied==1
# Step 2: Use select to keep the columns: Pclass, Age, SibSp, Parch, Fare
# Correlations on the passengers that survived


# Step 1: Use filter to keep observations with Survivied==0
# Step 2: Use select to keep the columns: Pclass, Age, SibSp, Parch, Fare
# Correlations on the passengers that did not survive


```


## Pairwise plot

Create a pairwise plot for variables including 
SurvivedLabel, Pclass, Sex, Age, SibSp, Parch, Fare, Embarked.

```{r,fig.width=12,fig.height=12}
library(GGally)

# Step 1: Convert Pclass to a factor
# Step 2: Use select to keep the above mentioned columns
# Step 3: do ggpairs()


```



# Feature Engineering on the titanicing Set {.tabset .tabset-fade}

What variable can we add?

## Discretize Age

Use the function `cut`.

```{r}
# create age_cut by cutting Age into c("children","adult","senior") 
summary(titanic$Age)
titanic <- titanic %>%
  mutate(age_bin=cut(Age, breaks = c(0,18,60,80),
                     labels=c("children","adult","senior"),
                     include.lowest = TRUE)) %>%
  mutate(age_bin=addNA(age_bin)) 
titanic %>% select(Age,age_bin) %>% print(n=2)

# Categories with NA (not available, i.e., missing values) added
levels(titanic$age_bin)

# Barplot of survival rate for each age_bin
titanic %>% 
  group_by(age_bin) %>%
  summarize(sur=mean(Survived)) %>%
  ggplot(aes(x=age_bin,y=sur))+
  geom_bar(stat="identity")+
  geom_text(aes(label=scales::percent(sur,
                                      accuracy=0.1,
                                      suffix="%") ),  
            size=3.5,vjust=-0.25,
            position = position_dodge(width = 0.9))+
  scale_y_continuous(labels = scales::percent, limits = c(0,1))+
  labs(y="Survival Rate")+
  theme_bw()
```


## Discretize Fare

Use the function `cut`.

```{r}
# create fare_cut by cutting Fare into c("low","medium","high")
# using the 1st quantile and the 3rd quantile


# Barplot of survival rate for each fare_bin



```

## Family size

Family size = SibSp + Parch

```{r}
# Add the variable FamilySize


# Use summarise and n() to count each FamilySize
# as well as mean() to compute the survival rate
# Step 1: group_by
# Step 2: summarise


# Bar chart for the surival rate for each FamilySize


```

## Single

According to the family size,
we can create a variable `single` which is TRUE if `FamilySize<1`.

```{r}
# Add the variable Single


# Use summarise and n() to count by Single
# as well as mean() to compute the survival rate
# Step 1: group_by
# Step 2: summarise

# Bar chart for the surival rate by Single


```


## Tickets

There are 891 obervations in this data set but
only 681 unique tikets.
So there were passengers with the same ticket.
This infomation might help predict the survival rate.

Since there are so many levels for this categorical variable,
we may use the count encoding for `Ticket`.

```{r}
# Create a new variable: shareTicket
# Step 1: Group by Ticket
# Step 2: summarise and n() to count the number of passengers for each ticket
# Step 3: Use mutate to create a varaible shareTicket which is TRUE if the ticket was shared by more than one person
# Step 4: Join to the original data set


# Use summarise and n() to count each shareTicket
# as well as mean() to compute the survival rate
# Step 1: group_by
# Step 2: summarise


# Bar chart for the surival rate for each shareTicket

```


## Target encoding 

Target encoding for Pclass is not useful because 
the Pclass has too few levels and 
one can apply the one-hot encoding without any difficulty.

In practice, we do not have the (Survived) labels in the testing set.
So we will use the what are learnt in the training set for the categories 
appear in testing set.

```{r}
# Create a new variable: surPclass
# Step 1: Group by Pclass
# Step 2: summarise and mean to find the survival ratio for each Pclass
# Step 3: Join to the original data set
# Step 4: Add random noises rnorm(nrow(titanic),sd=0.01) to the 

```

# Predict Survival Rate

We use the logistic regression to model the probability (rate) of survival.
See "Titanic_solution.Rmd" for details.









