---
title: "Exploratory Data Analysis on Titanic Surivival"
author: "Your name"
date: "`r Sys.Date()`"
output:
 bookdown::gitbook:
  includes:
    in_header: header.html
  self_contained: false  
  split_by: none
  config:
    toc:
      collapse: subsubsection
      scroll_highlight: yes
    toolbar:
      position: fixed
    search: yes
    fontsettings:
      theme: white
      family: sans
      size: 3
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
```

```{r header, echo=FALSE}
# Do not modify
codejs <- readr::read_lines("js/codefolding.js")
collapsejs <- readr::read_lines("js/collapse.js")
transitionjs <- readr::read_lines("js/transition.js")
dropdownjs <- readr::read_lines("js/dropdown.js")

htmlhead <- c(
  paste('
<script>',
paste(transitionjs, collapse = "\n"),
'</script>
<script>',
paste(collapsejs, collapse = "\n"),
'</script>
<script>',
paste(codejs, collapse = "\n"),
'</script>
<script>',
paste(dropdownjs, collapse = "\n"),
'</script>
<style type="text/css">
.code-folding-btn { margin-bottom: 4px; }
.row { display: flex; }
.collapse { display: none; }
.in { display:block }
.pull-right > .dropdown-menu {
    right: 0;
    left: auto;
}
.open > .dropdown-menu {
    display: block;
}
.dropdown-menu {
    position: absolute;
    top: 100%;
    left: 0;
    z-index: 1000;
    display: none;
    float: left;
    min-width: 160px;
    padding: 5px 0;
    margin: 2px 0 0;
    font-size: 14px;
    text-align: left;
    list-style: none;
    background-color: #fff;
    -webkit-background-clip: padding-box;
    background-clip: padding-box;
    border: 1px solid #ccc;
    border: 1px solid rgba(0,0,0,.15);
    border-radius: 4px;
    -webkit-box-shadow: 0 6px 12px rgba(0,0,0,.175);
    box-shadow: 0 6px 12px rgba(0,0,0,.175);
}
</style>
<script>
$(document).ready(function () {
  window.initializeCodeFolding("show" === "hide");
});
</script>
', sep = "\n"),
  paste0('
<script>
document.write(\'<div class="btn-group pull-right" style="position: absolute; top: 20%; right: 2%; z-index: 200"><button type="button" class="btn btn-default btn-xs dropdown-toggle" data-toggle="dropdown" aria-haspopup="true" aria-expanded="true" data-_extension-text-contrast=""><span>Code</span> <span class="caret"></span></button><ul class="dropdown-menu" style="min-width: 50px;"><li><a id="rmd-show-all-code" href="#">Show All Code</a></li><li><a id="rmd-hide-all-code" href="#">Hide All Code</a></li></ul></div>\')
</script>
')
)

readr::write_lines(htmlhead, path = "header.html")
```

# Preparations {.tabset .tabset-fade}

## Load libraries

```{r, message=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(scales)
setwd("C:/Users/Yi/Dropbox/newtermphd/Term12/Rworkshop_data_manipulation/exercise/Titanic/gitbook/") 
```


## Read data
```{r data}
titanic <- read_csv("titanic.csv")
```

# Glance at data  {.tabset .tabset-fade}

## Data structure
```{r data str,result='asis', echo=FALSE}
titanic
titanic %>% glimpse()
```


## Number of distinct values for each variable

Make use of `n_distinct`

```{r}
# Count by summarize_all
titanic %>%
  summarize_all(n_distinct) %>%
  gather(key = "variable",value="Unique")
```


## Missing rate for each variable

```{r}
# Count by summarize_all
titanic %>%
  summarize_all(list(~sum(is.na(.))/n())) %>%
  gather(key = "variable",value="missingRate")

# Create a bar chart to display it
titanic %>%
  summarize_all(list(~sum(is.na(.))/n())) %>%
  gather(key = "variable",value="missingRate") %>%
  ggplot(aes(variable,missingRate))+
  geom_bar(stat="identity",fill="red")+
  theme_bw()
```

## Create a variable SurvivedLabel as a factor of Survival

```{r}
titanic

# Create a variable SurvivedLabel as a factor of Survival
titanic <- titanic %>%
  mutate(SurvivedLabel=factor(Survived))

```



# Survival for each variable  {.tabset .tabset-fade}

## Overall

Count the number and proportion of survived passengers.

```{r}
# Step 1: group_by SurvivedLabel
# Step 2: summarise count using count=n()
# Step 3: mutate to find the proportion, count/sum(count)
titanic %>% 
  group_by(SurvivedLabel) %>%
  summarise(count=n()) %>%
  mutate(prop=count/sum(count))
```


## Pclass

Ticket class.

```{r}
# Use summarise and n() to count each Pclass
# as well as mean() to compute the survival rate
# Step 1: group_by
# Step 2: summarise
titanic %>% 
  group_by(Pclass) %>%
  summarise(sur=mean(Survived), 
            count=n())

# Bar chart for the surival rate
titanic %>% 
  group_by(Pclass) %>%
  summarise(sur=mean(Survived)) %>%
  ggplot(aes(Pclass,sur)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=scales::percent(sur,
                                      accuracy=0.1,
                                      suffix="%") ),  
            size=3.5,vjust=-0.25,
            position = position_dodge(width = 0.9))+
  scale_y_continuous(labels = scales::percent, limits = c(0,1))+
  labs(y="Survival Rate")+
  theme_bw()

```


## Sex

```{r}
# Use summarise and n() to count each Sex
# as well as mean() to compute the survival rate
# Step 1: group_by
# Step 2: summarise
titanic %>% 
  group_by(Sex) %>%
  summarise(sur=mean(Survived), 
            count=n())

# Bar chart for the surival rate
titanic %>% 
  group_by(Sex) %>%
  summarise(sur=mean(Survived)) %>%
  ggplot(aes(Sex,sur)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=scales::percent(sur,
                                      accuracy=0.1,
                                      suffix="%") ),  
            size=3.5,vjust=-0.25,
            position = position_dodge(width = 0.9))+
  scale_y_continuous(labels = scales::percent, limits = c(0,1))+
  labs(y="Survival Rate")+
  theme_bw()

```

## Age

Use `SurvivedLabel` as a group indicator in boxplot.
Notice that there are many missing values.

```{r}
# Create a histogram of Age
titanic %>% 
  ggplot(aes(x=Age))+
  geom_histogram()

# Boxplot between SurvivedLabel and Age
titanic %>% 
  ggplot(aes(SurvivedLabel,Age)) +
  geom_boxplot()

```

## SibSp

Number of siblings / spouses aboard the Titanic.
The survival rate within each group indicate that 
the survival rate drops for people with higher SibSp except for people who had no siblings.

```{r}
# Use summarise and n() to count each SibSp
# as well as mean() to compute the survival rate
# Step 1: group_by
# Step 2: summarise
titanic %>% 
  group_by(SibSp) %>%
  summarise(sur=mean(Survived), 
            count=n())

```


## Parch

Number of parents / children aboard the Titanic.

```{r}
# Use summarise and n() to count each Parch
# as well as mean() to compute the survival rate
# Step 1: group_by
# Step 2: summarise
titanic %>% 
  group_by(Parch) %>%
  summarise(sur=mean(Survived), 
            count=n())

```


## Fare

Passenger fare.

```{r}
# Boxplot between SurvivedLabel and Fare
titanic %>% 
  ggplot(aes(SurvivedLabel,Fare)) +
  geom_boxplot()

# Boxplot between factor(Pclass) and Fare
titanic %>% 
  ggplot(aes(factor(Pclass),Fare)) +
  geom_boxplot()

```

## Embarked

Port of Embarkation.

```{r}
# Use summarise and n() to count each Embarked
# as well as mean() to compute the survival rate
# Step 1: group_by
# Step 2: summarise
titanic %>% 
  group_by(Embarked) %>%
  summarise(sur=mean(Survived), 
            count=n())

# Bar chart for the surival rate
titanic %>% 
  group_by(Embarked) %>%
  summarise(sur=mean(Survived)) %>%
  ggplot(aes(Embarked,sur)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=scales::percent(sur,
                                      accuracy=0.1,
                                      suffix="%") ),  
            size=3.5,vjust=-0.25,
            position = position_dodge(width = 0.9))+
  scale_y_continuous(labels = scales::percent, limits = c(0,1))+
  labs(y="Survival Rate")+
  theme_bw()

```


## Sex and Pclass

```{r}
# Survival rates and counts for the combinations of Sex and Pclass
# Use mutate(SexPclass=paste0(Sex,"_",Pclass))
# to create an interaction variable
titanic %>% 
  group_by(Sex,Pclass) %>%
  summarise(sur=mean(Survived),
            count=n()) %>%
  ungroup() %>%
  mutate(SexPclass=paste0(Sex,"_",Pclass))

# Bar chart for the surival rate
titanic %>% 
  group_by(Sex,Pclass) %>%
  summarise(sur=mean(Survived)) %>%
  ungroup() %>%
  mutate(SexPclass=paste0(Sex,"_",Pclass)) %>%
  ggplot(aes(SexPclass,sur)) +
  geom_bar(stat="identity")

```

# Others {.tabset .tabset-fade}

## Correlation

Find Spearman Rank's correlations among
Survived, Pclass, Age, SibSp, Parch, Fare.
Hint: Use `cor(use="complete.obs", method = "spearman")`.

```{r}
library(corrplot)

# Step 1: Use select to keep the columns: Survived, Pclass, Age, SibSp, Parch, Fare
# Correlations on all the passengers
titanic %>% 
  select(Survived,Pclass,Age, SibSp, Parch, Fare) %>%
  cor(use="complete.obs", method = "spearman") %>%
  corrplot.mixed(main="Correlations on all the passengers",
                 mar = c(0, 0, 2, 0))

# Step 1: Use filter to keep observations with Survivied==1
# Step 2: Use select to keep the columns: Pclass, Age, SibSp, Parch, Fare
# Correlations on the passengers that survived
titanic %>% 
  filter(Survived==1) %>%
  select(Pclass, Age, SibSp, Parch, Fare) %>%
  cor(use="complete.obs", method = "spearman") %>%
  corrplot.mixed(main="Correlations on the passengers that survived",
                 mar = c(0, 0, 2, 0))

# Step 1: Use filter to keep observations with Survivied==0
# Step 2: Use select to keep the columns: Pclass, Age, SibSp, Parch, Fare
# Correlations on the passengers that did not survive
titanic %>% 
  filter(Survived==0) %>%
  select(Pclass, Age, SibSp, Parch, Fare) %>%
  cor(use="complete.obs", method = "spearman") %>%
  corrplot.mixed(main="Correlations on the passengers that did not survive",
                 mar = c(0, 0, 2, 0))
```


## Pairwise plot

Create a pairwise plot for variables including 
SurvivedLabel, Pclass, Sex, Age, SibSp, Parch, Fare, Embarked.

```{r,fig.width=12,fig.height=12}
library(GGally)

# Step 1: Convert Pclass to a factor
# Step 2: Use select to keep the above mentioned columns
# Step 3: do ggpairs()

titanic %>% 
  mutate(Pclass=factor(Pclass)) %>%
  select(SurvivedLabel, Pclass, Sex, Age, 
         SibSp, Parch, Fare, Embarked) %>%
  ggpairs()
```



# Feature Engineering on the titanicing Set {.tabset .tabset-fade}

What variable can we add?

## Discretize Age

Use the function `cut`.

```{r}
# create age_cut by cutting Age into c("children","adult","senior") 
summary(titanic$Age)
titanic <- titanic %>%
  mutate(age_bin=cut(Age, breaks = c(0,18,60,80),
                     labels=c("children","adult","senior"),
                     include.lowest = TRUE)) %>%
  mutate(age_bin=addNA(age_bin)) 
titanic %>% select(Age,age_bin) %>% print(n=2)

# Categories with NA (not available, i.e., missing values) added
levels(titanic$age_bin)

# Barplot of survival rate for each age_bin
titanic %>% 
  group_by(age_bin) %>%
  summarize(sur=mean(Survived)) %>%
  ggplot(aes(x=age_bin,y=sur))+
  geom_bar(stat="identity")+
  geom_text(aes(label=scales::percent(sur,
                                      accuracy=0.1,
                                      suffix="%") ),  
            size=3.5,vjust=-0.25,
            position = position_dodge(width = 0.9))+
  scale_y_continuous(labels = scales::percent, limits = c(0,1))+
  labs(y="Survival Rate")+
  theme_bw()
```


## Discretize Fare

Use the function `cut`.

```{r}
# create fare_cut by cutting Fare into c("low","medium","high")
# using the first quantile and the third quantile
summary(titanic$Fare)
titanic <- titanic %>%
  mutate(fare_bin=cut(Fare, breaks = c(0,8,31,550),
                     labels=c("low","medium","high"),
                     include.lowest = TRUE))
  
titanic %>% select(Fare,fare_bin) %>% print(n=5)


# Barplot of survival rate for each fare_bin
titanic %>% 
  group_by(fare_bin) %>%
  summarize(sur=mean(Survived)) %>%
  ggplot(aes(x=fare_bin,y=sur))+
  geom_bar(stat="identity")+
  geom_text(aes(label=scales::percent(sur,
                                      accuracy=0.1,
                                      suffix="%") ),  
            size=3.5,vjust=-0.25,
            position = position_dodge(width = 0.9))+
  scale_y_continuous(labels = scales::percent, limits = c(0,1))+
  labs(y="Survival Rate")+
  theme_bw()
```

## Family size

Family size = SibSp + Parch

```{r}
# Add the variable FamilySize
titanic <- titanic %>%
  mutate(FamilySize=SibSp + Parch)


# Use summarise and n() to count each FamilySize
# as well as mean() to compute the survival rate
# Step 1: group_by
# Step 2: summarise
titanic %>% 
  group_by(FamilySize) %>%
  summarise(sur=mean(Survived), 
            count=n())

# Bar chart for the surival rate for each FamilySize
titanic %>% 
  group_by(FamilySize) %>%
  summarise(sur=mean(Survived)) %>%
  ggplot(aes(FamilySize,sur)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=scales::percent(sur,
                                      accuracy=0.1,
                                      suffix="%") ),  
            size=3.5,vjust=-0.25,
            position = position_dodge(width = 0.9))+
  scale_y_continuous(labels = scales::percent, limits = c(0,1))+
  labs(y="Survival Rate")+
  theme_bw()


```

## Single

According to the family size,
we can create a variable `single` which is TRUE if `FamilySize<1`.

```{r}
# Add the variable Single
titanic <- titanic %>%
  mutate(Single= FamilySize<1)

# Use summarise and n() to count by Single
# as well as mean() to compute the survival rate
# Step 1: group_by
# Step 2: summarise
titanic %>% 
  group_by(Single) %>%
  summarise(sur=mean(Survived), 
            count=n())

# Bar chart for the surival rate by Single
titanic %>% 
  group_by(Single) %>%
  summarise(sur=mean(Survived)) %>%
  ggplot(aes(Single,sur)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=scales::percent(sur,
                                      accuracy=0.1,
                                      suffix="%") ),  
            size=3.5,vjust=-0.25,
            position = position_dodge(width = 0.9))+
  scale_y_continuous(labels = scales::percent, limits = c(0,1))+
  labs(y="Survival Rate")+
  theme_bw()
```


## Tickets

There are 891 obervations in this data set but
only 681 unique tikets.
So there were passengers with the same ticket.
This infomation might help predict the survival rate.

Since there are so many levels for this categorical variable,
we may use the count encoding for `Ticket`.

```{r}
# Create a new variable: shareTicket
# Step 1: Group by Ticket
# Step 2: summarise and n() to count the number of passengers for each ticket
# Step 3: Use mutate to create a varaible shareTicket which is TRUE if the ticket was shared by more than one person
# Step 4: Join to the original data set
titanic <- titanic %>% 
  group_by(Ticket) %>%
  summarise(shareTicket=n()) %>% 
  mutate(shareTicket=shareTicket>1) %>%
  right_join(titanic, by ="Ticket")


# Use summarise and n() to count each shareTicket
# as well as mean() to compute the survival rate
# Step 1: group_by
# Step 2: summarise
titanic %>% 
  group_by(shareTicket) %>%
  summarise(sur=mean(Survived), 
            count=n())

# Bar chart for the surival rate for each shareTicket
titanic %>% 
  group_by(shareTicket) %>%
  summarise(sur=mean(Survived)) %>%
  ggplot(aes(shareTicket,sur)) +
  geom_bar(stat="identity")+
  geom_text(aes(label=scales::percent(sur,
                                      accuracy=0.1,
                                      suffix="%") ),  
            size=3.5,vjust=-0.25,
            position = position_dodge(width = 0.9))+
  scale_y_continuous(labels = scales::percent, limits = c(0,1))+
  labs(y="Survival Rate")+
  theme_bw()

```


## Target encoding 

Target encoding for Pclass is not useful because 
the Pclass has too few levels and 
one can apply the one-hot encoding without any difficulty.

In practice, we do not have the (Survived) labels in the testing set.
So we will use the what are learnt in the training set for the categories 
appear in testing set.

```{r}
# Create a new variable: surPclass
# Step 1: Group by Pclass
# Step 2: summarise and mean to find the survival ratio for each Pclass
# Step 3: Join to the original data set
# Step 4: Add random noises rnorm(nrow(titanic),sd=0.01) to the 
titanic %>% 
  group_by(Pclass) %>%
  summarise(surPclass=mean(Survived)) %>%
  right_join(titanic, by ="Pclass") %>%
  select(Pclass,surPclass) %>%
  mutate(surPclass=surPclass+rnorm(nrow(titanic),sd=0.01))        # add noise

```

# Predict Survival Rate

We use the logistic regression to model the probability (rate) of survival.

Data are split into the training set and the testing set.


```{r}
# Convert categories to factors for variables:  Pclass, Sex, Single, Embarked
# filter out missing values of Embarked
# select columns: SurvivedLabel, Pclass, Sex, age_bin, fare_bin, Embarked
# Assign to a new R object, titanic2
titanic2 <- titanic %>%
  mutate(Pclass=factor(Pclass), Sex=factor(Sex), 
         Single=factor(Single),Embarked=factor(Embarked)) %>%
  filter(!is.na(Embarked)) %>%
  select(SurvivedLabel, Pclass, Sex, Single,age_bin, fare_bin, Embarked)

titanic2

# Randomly split into train and test
set.seed(111)
train_idx <- sample(1:nrow(titanic2), size = 500, replace = FALSE)
train <- titanic2[train_idx,]
test <- titanic2[-train_idx,]

# Train logisitc regression model
fit1 <- glm(SurvivedLabel~Pclass+Sex+Single+age_bin+fare_bin+Embarked,
            data=train, family = binomial)

# Estimated coefficients and significances
summary(fit1)

# Perform backward stepwise model selection according to the AIC (smaller is better)
fit2 <- step(fit1)
summary(fit2) #Single, fare_bin is removed

# Predict the survival rates for passengers in the testing set
pred_fit2 <- predict(fit2, newdata = test, 
                     type = "response", se.fit = TRUE)
# Format the predictions
format_pred <- tibble(predSurv=ifelse(pred_fit2$fit>0.5,1,0),
                      survRate=pred_fit2$fit,
                      se=pred_fit2$se.fit)
format_pred

# Accuracy
mean(format_pred$predSurv == test$SurvivedLabel)

# Confusion table
table(Prediction=format_pred$predSurv, Truth=test$SurvivedLabel)


#------------- Forecasts for all combinations ------------------#
newtest <- expand.grid( Pclass = c("1","2","3"),
                        Sex = c("male","female"),
                        age_bin = c("children","adult","senior",NA),
                        Embarked = c("C","Q","S"))
pred_newtest <- predict(fit2, newdata = newtest, 
                     type = "response")

# Format
as_tibble(newtest) %>% 
  bind_cols(tibble(survRate=pred_newtest)) %>%
  arrange(desc(survRate)) %>%
  print(n=100)

```




